Global optimization: from one-dimensional to multidimensional problems via the dimensionality reduction

### 2017-10-07 2017/ParOptBook/Chapter1.tex
Chapter -> paper

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%
%% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

@@

\abstract{This chapter is devoted to the basic statements and general definitions concerning the global optimization problems and algorithms. At the beginning, relevant notions are introduced and different statements of global optimization problems are described. As a main optimization statement, a non-linear programming problem is considered.  The concepts of numerical methods for solving the optimization problems are discussed and a general abstract model for successive methods is introduced. On the base of abstract description its algorithmic implementation (general computational scheme) is given. The model of sequential optimization expands for the case of parallel computations and general computational procedures for synchronous and asynchronous parallel methods are presented. Convergence of optimization methods and possibility of obtaining estimations of the problem solution after a finite number of objective function evaluations are discussed. Significant complexity of multidimensional multiextremal optimization problems is considered and approaches of overcoming this complexity via reduction schemes are briefly described. }

@@

Let us designate 
\begin{equation}
\label{eq:1}
\inf_{y\in Q}\varphi(y) \equiv\inf\{\varphi(y):y \in Q\}
\end{equation}
as the infimum of $\varphi(y)$ in the domain $Q$.
\begin{definition}
If there exists a point  $y^*\in Q$  such that 
\begin{equation}
\label{eq:2}
\varphi(y^*)=\inf\{\varphi(y):y \in Q\}
\end{equation}
then $\varphi(y)$ reaches its infimum in the domain $Q$, and $y^*$ is \textit {the global (absolute) minimum point} or \textit {the global minimizer}. The value $\varphi^* = \varphi(y^*)$ is the lowest or global optimum (minimum) value of $\varphi$ in $Q$ and is denoted as 
\begin{equation}
\label{eq:3}
\min_{y\in Q}\varphi(y) \equiv\min\{\varphi(y):y \in Q\}.
\end{equation}
\end{definition}
Let us denote the set of all points $y\in Q$  satisfying (\ref{eq:2}) as
\begin{equation}
\label{eq:4}
Q^*=Arg\min_{y\in Q}\varphi(y) \equiv Arg\min\{\varphi(y):y \in Q\}.
\end{equation}

@@

\textbf{Problem Statement A.} Find the infimum of the function $\varphi(y)$
\begin{equation}
\label{eq:5}
\varphi^*=\inf\{\varphi(y):y \in Q\}.
\end{equation}

\textbf{Problem Statement B.} Find the infimum from (\ref{eq:5}) and if the set of the global minimum points from (\ref{eq:2}) is not empty, find at least one point $y^*=Q^*$.

\textbf{Problem Statement C.} Find the infimum from (\ref{eq:5})  and find all global minimum points (or make sure that the set $Q^*$ is empty).

@@

The problem statements A—D are related to the finding of the global minimum, but sometimes the following problem statement is used even for the multiextremal functions:

@@

The optimization problem statements A–E are the most commonly used (see, for example, classical monographs \cite{1_FloudasPardalos, 1_Himmelblau, 1_HorstPardalos, 1_HorstTuy, 1_McCormick, 1_Mockus, 1_Pinter,  1_SergKvasMonogr, 1_StrSergMon2000, 1_ZhigZhil}), although other variants are possible as well, for example, those given in \cite{1_Himmelblau}. The desired extremal characteristics defined by the problem statement $Z\in\{A, B, C, D, E\}$ are referred to as $Z$-solution of the optimization problem.

Let us denote the general optimization problem symbolically in the following form:
\begin{equation}
\label{eq:6}
\varphi(y)\rightarrow\inf, y \in Q.
\end{equation}

@@

As  the supremum of the function $\varphi(y)$ in the domain $Q$ is connected with the infimum of the function $-\varphi(y)$ according to  the expression
\begin{displaymath}
\sup\{\varphi(y):y \in Q\} = -\inf\{-\varphi(y):y \in Q\}
\end{displaymath}
the problem of finding the extremal characteristics related to the maximum of the function $\varphi(y)$ (the maximization problem), can be reduced to the problem of the minimization of the function $-\varphi(y)$. Therefore, hereinafter the optimization problem will be considered in the form (\ref{eq:6}). The function $-\varphi(y)$ in (\ref{eq:6}) will be called \textit{the objective function, the minimized function, or the optimized function},  the domain $Q$ will be called \textit{the feasible (admissible) domain}, and the elements of the set $Q$ will be called \textit{the feasible points}.

Let us formulate the problem (\ref{eq:6}), for which the set of the global minimum points $Q^*$ is known to be nonempty a priori (the sufficient conditions for the non-emptiness of $Q^*$ are given, for example, by Weierstrass theorem), as
\begin{equation}
\label{eq:7}
\varphi(y)\rightarrow\min, y \in Q.
\end{equation}

Note that the mathematical programming problem is often formulated just in this form, although, of course, the problem (\ref{eq:6}) is the more general one.
In the present monograph, the existence of a global minimizer is always presumed. Let us reformulate the problem (\ref{eq:8}) in the form of a non-linear programming problem
\begin{equation}
\label{eq:8}
\varphi(y)\rightarrow\min, y \in Q\subseteq R^N,
\end{equation}
\begin{equation}
\label{eq:9}
Q=\{y \in D:g_j(y)\leq 0,1\leq j\leq m \},
\end{equation}
\begin{equation}
\label{eq:10}
D=\{y \in R^N:y_i\in [a_i,b_i],1\leq i\leq N \},
\end{equation}
i.e., the problem of finding the extremal values (in terms of the Problem statements A-D) of the objective (minimized) function $\varphi(y)$  over the domain $Q$ is defined by the \textit{coordinate} (\ref{eq:10}) and \textit{functional} (\ref{eq:9}) \textit{constraints} on the selection of feasible points (or vectors) $y=(y_1,\ldots ,y_N)$ . In this model, the constants $a_i,b_i, 1\leq i\leq N$,  determine the limits for the variation of the problem parameters (the components of the vector $y$), and are finite or (when the corresponding upper or/and lower limits are absent) are assumed to be equal to $a_i=-\infty $  and/or $b_i=+\infty $ .

Often, the constraints in the form of equations are also included in the non-linear programming problem statement. However, any equation $h(y)=0$ , first, could be represented formally as the system of two inequalities $h(y)\leq 0$  and $h(y)\geq 0$ . Second, in numerical solving an optimization problem on a computer the exact calculation of the equality $h(y)=0$ is often impossible, therefore, this equality is assumed to be computed  with some error $\delta>0$ , i.e., the inequality $\left|h(y)\right|\leq \delta$ is considered instead of the equation $h(y)=0$. Thus, taking into account the above reasons, one can claim that (\ref{eq:8})--(\ref{eq:10}) is the statement of  \textit {the general  problem of the nonlinear programming}.

If $m=0$ , i.e., there are no functional constraints, let us assume $Q=D$. The problem (\ref{eq:8})--(\ref{eq:10}) in this case will be called \textit {the unconstrained optimization problem}.

Let us consider an optimization problem  for the function
\begin{displaymath}
%\label{eq:12}
\varphi(y_1,y_2)=-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)-
\end{displaymath}
\begin{displaymath}
%\label{eq:12}
-(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
\end{displaymath}
within the domain $0 \leq y_1\leq 4,\ \  -1\leq y_2\leq 3$, with the following constraints 
\begin{displaymath}
g_1(y_1,y_2)=0.01((y_1-2.2)^2+(y_2-1.2)^2-2.25)\leq 0,
\end{displaymath}
\begin{displaymath}
g_2(y_1,y_2)=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,
\end{displaymath}
\begin{displaymath}
g_3(y_1,y_2)=10(y_2-1.5-1.5\sin(6.283(y_1-1.75)))\leq 0,
\end{displaymath}
as a particular example.

@@

The set of the points satisfying the first constraint represents a circle with the boundary defined by the equation $g_1(y_1,y_2)=0$. The points satisfying the second constraint are located outside an ellipse defined by the equation $g_2(y_1,y_2)=0$. The points satisfying the third constraint are below the sinusoid $g_3(y_1,y_2)=0$. Thus, the feasible domain is disconnect and consists of three non-convex subdomains (the feasible domain is dark in Fig.~\ref{1_fig_1}). The global minimum $\varphi(y_1^*,y_2^*)=-1.489$  is achieved at the point $(y_1^*,y_2^*)$  = (0.942, 0.944).

@@

Unfortunately, these requirements are fulfilled rarely. In the typical cases, the function $\varphi(y)$ is defined algorithmically, i.e., in the form of some calculation scheme, when one can calculate the value of $\varphi(y)$ at a given point y only (black-box function). In this case, the analytical investigation methods are unsuitable. Note that even if the function is defined analytically and if the derivative calculation is possible the initial problem (\ref{eq:6}) is reduced to the problem of finding all the derivative roots, the difficulty of which is comparable with the minimization problem.

@@

Let us give a formal definition of sequential optimization algorithm based on the methodology of the theory of operations research, or, more general, on the computational model for solving the problem (\ref{eq:6}). The generalization of this model onto the parallel computing case will be given in Subsect. \ref{subsec:1.2.2}

@@

The selection of an \textit {optimization algorithm} (method for solving  the optimization problem) is the next important step in the computational model development. In the most general form, a numerical method $s$ for solving the problem from the class $\Phi$ is a set  (tuple) (\cite{1_StrMonRus})
\begin{equation}
\label{eq:11}
s=\left\langle \{G_k\},\{E_k\},\{H_k\}\right\rangle
\end{equation}
where

@@

-  $\{E_k\}$	  is a set of functionals defining the rules of constructing the approximate solution (the estimate of extremum), $k=1,2,\ldots$ ;

@@

\begin{description}[\textbf{Step 1}]
\item[\textbf{Step 1}]{The point 
\begin{equation}
\label{eq:12}
y^1=G_1(\Phi\in Q)
\end{equation}
is selected as the first trial coordinate and the trial counter $k=1$ .}
\item[\textbf{Step 2}]{Assume the point of the $k$-th trial $y^k\in Q$  $(k\geq 1)$ has been selected. The value of the function $z^k=\varphi(y^k)$  is computed. Afterwards there is the following \textit {search (a posteriori) information} about the function $\varphi(y)$:
\begin{equation}
\label{eq:13}
\omega_k=\{(y^1,z^1),(y^2,z^2),\ldots ,(y^k,z^k)\}.
\end{equation}
This information allows reducing the class, which the function  $\varphi(y)$ belongs to, up to the set
\begin{equation}
\label{eq:14}
\Phi(\omega_k)=\{\psi\in\Phi:\psi(y^i)=z^i,1\leq i \leq k\}.
\end{equation}
}
\item[\textbf{Step 3}]{The current estimate 
\begin{equation}
\label{eq:15}
e^k=E_k(\Phi,\omega_k)
\end{equation}
of the extremum sought (approximate solution) is determined.}
\item[\textbf{Step 4}]{The next trial point 
\begin{equation}
\label{eq:16}
y^{k+1}=G_{k+1}(\Phi,\omega_k)
\end{equation}
is computed.}
\item[\textbf{Step 5}]{The number 
\begin{equation}
\label{eq:17}
h^k=H_k(\Phi,\omega_k,y^{k+1})\in \{0,1\}
\end{equation}
taking one of two possible values: 0 or 1 is determined. If $h^k=1$ the trial counter $k$ is increased by 1, and the return to the Step 2 of the scheme is carried out. If $h^k=0$ the computations are terminated, and the estimate $e^k$  is taken as the problem solution.}
\end{description}

@@

\example {Let us consider the simplest method of solving a one-dimensional problem (\ref{eq:6}) within an interval $[a,b]$, namely, the method of scanning  the function values in the nodes of a regular grid. The method consists in the partitioning of the search interval $[a,b]$ into $q$ equal parts. The function values are calculated at the points (nodes) of the partitioning, including the ends of the search interval. The lowest calculated value (and its coordinate, if it is required by the problem statement) is considered to be the solution of the problem. }

The computability of the function at any point of the search domain is sufficient for the applicability of the method. So, the class of the functions defined over the interval $[a,b]$ and being computable at every point of the interval could be considered as the a priori known class $\Phi$.

For this method
\begin{displaymath}
G_1(\Phi)=a, G_{k=1}(\Phi,\omega_k)=a+k\frac{b-a}{q}, k\geq 1,
\end{displaymath}
\begin{displaymath}
  H_k(\Phi,\omega_k) =
  \begin{cases}
    0, & k > q, \\
    1, & k \leq q,
  \end{cases}
\end{displaymath}
\begin{equation}
\label{eq:18}
e^k=\varphi_k^*,
\end{equation}
where
\begin{equation}
\label{eq:19}
\varphi_k^*=\min_{1\leq i\leq k} \varphi(y^i),
\end{equation}
or
\begin{displaymath}
e^k=(\varphi_k^*,y_k^*),
\end{displaymath}
where 
\begin{displaymath}
%\label{eq:19}
y_k^*=\arg \min_{1\leq i\leq k} \varphi(y^i).
\end{displaymath}

\subsection {Parallel optimization algorithms}
\label {subsec:1.2.2}
The computational scheme (\ref{eq:12})--(\ref{eq:17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers. But if a multiprocessor computer is available, a parallelization of the computational scheme and, therefore, an acceleration of the optimization process can be realized. The parallelization of optimization methods could be done in several ways. It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:12})--(\ref{eq:17}) providing simultaneous execution of several trials. However, the parallelization of the objective function describing the optimized object is specific for different objective function models, and as a matter of fact, cannot be related to the general optimization scheme. The parallelization of the functionals  $G_k,E_k,H_k$ depends essentially on the class of algorithms. Besides, often these functionals are rather simple, and their parallelization is just of no use. Therefore, Therefore, the subject of the consideration in the present subsection  is the design of some \textit {fundamental principles} for a parallel execution of trials.  Let us use the term \textit {iteration} for the simultaneous (parallel) execution of several trials (each trial is executed by a separate processor). Let us denote the number of trials during the $n$-th iteration as $p(n)$ , and the total number of trials executed during all $n$ iterations as $k(n)$ . In other words, we assume that we have  $p$  processors at our disposal while executing the $n$-th iteration. Actually, the number of processors employed can be different for different iterations.

@@

Thereby, let us ask the question: what requirements should be imposed to trial sequences of a numerical optimization method? Of course, the main requirement is that execution of the trials at the points $\{y^k\}$  with the trial results $\{z^k\}$  could provide a solution of the problem, i.e., finding the solution corresponding to the selected problem statement. At the same time, since a computer can execute a finite number of trials only, it is desirable to obtain the exact solution after constructing a finite sequence $\{y^k\}$. Therefore,  an asymptotically exact assessment is interesting often, and an infinite trial sequence is considered (in the introduced  optimization method models in (\ref{eq:17}) , (\ref{eq:25}), (\ref{eq:30}) $H_k=1$  for any $k\geq1$, i.e., in fact, the termination criterion is absent). This sequence is required to converge to the exact problem solution. Since in the problem statements $B-D$ the desired solution may include several minimum points, we will understand the convergence of a method in the scope of the following definition.
\begin{definition} 
\label{def:1_4}
The trial sequence $\{y^k\}$  converges to the solution of the problem (\ref{eq:6}) defined by the corresponding problem statement, if:
\begin{enumerate}
\item{it includes a subsequence  $\{\bar{y}^k\}$, for which $\lim_{k \to \infty}\bar{y}^k=\varphi^*$;}
\item{in the case when the solution includes one or several minimum points, for each such point there exists a subsequence of the sequence $\{y^k\}$  converging to it.}
\end{enumerate}
\end{definition}
Trial sequence converging to the exact solution of the problem statement $Z \in \{A,B,C,D,E\}$ is called the minimizing sequence for the problem statement $Z$ or $Z$-minimizing sequence. The term «minimizing sequence» was introduced in \cite{1_Vasiliev} and corresponds to the concept of $A$-minimizing sequence.

@@

Another approach to the design of the numerical methods for the analysis of the multidimensional multiextremal problems exploits the idea of \textit{the complexity reduction} where the solution of the initial problem is replaced by solving one or several simpler problems.

>>>>

%%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% sample root file for your "contribution" to a contributed volume
%%
%% Use this file as a template for your own input.
%%
%%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\documentclass[graybox]{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
%\usepackage{mathptmx}       % selects Times Roman as basic font
%\usepackage{helvet}         % selects Helvetica as sans-serif font
%\usepackage{courier}        % selects Courier as typewriter font
%\usepackage{type1cm}        % activate if the above 3 fonts are
                            %% not available on your system
%%
%\usepackage{makeidx}         % allows index generation
%\usepackage{graphicx}        % standard LaTeX graphics tool
                             %% when including figure files
%\usepackage{multicol}        % used for the two-column index
%\usepackage[bottom]{footmisc}
%\usepackage{amsmath}
%
%% places footnotes at page bottom
%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
                       %% please use the style svind.ist with
                       %% your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%\begin{document}

@@

\abstract{This paper is devoted to the basic statements and general definitions concerning the global optimization problems and algorithms. At the beginning, relevant notions are introduced and different statements of global optimization problems are described. As a main optimization statement, a non-linear programming problem is considered.  The concepts of numerical methods for solving the optimization problems are discussed and a general abstract model for successive methods is introduced. On the base of abstract description its algorithmic implementation (general computational scheme) is given. The model of sequential optimization expands for the case of parallel computations and general computational procedures for synchronous and asynchronous parallel methods are presented. Convergence of optimization methods and possibility of obtaining estimations of the problem solution after a finite number of objective function evaluations are discussed. Significant complexity of multidimensional multiextremal optimization problems is considered and approaches of overcoming this complexity via reduction schemes are briefly described. }

@@

Let us designate 
\begin{equation}
\label{eq:1_1_1}
\inf_{y\in Q}\varphi(y) \equiv\inf\{\varphi(y):y \in Q\}
\end{equation}
as the infimum of $\varphi(y)$ in the domain $Q$.
\begin{definition}
If there exists a point  $y^*\in Q$  such that 
\begin{equation}
\label{eq:1_1_2}
\varphi(y^*)=\inf\{\varphi(y):y \in Q\}
\end{equation}
then $\varphi(y)$ reaches its infimum in the domain $Q$, and $y^*$ is \textit {the global (absolute) minimum point} or \textit {the global minimizer}. The value $\varphi^* = \varphi(y^*)$ is the lowest or global optimum (minimum) value of $\varphi$ in $Q$ and is denoted as 
\begin{equation}
\label{eq:1_1_3}
\min_{y\in Q}\varphi(y) \equiv\min\{\varphi(y):y \in Q\}.
\end{equation}
\end{definition}
Let us denote the set of all points $y\in Q$  satisfying (\ref{eq:1_1_2}) as
\begin{equation}
\label{eq:1_1_4}
Q^*=Arg\min_{y\in Q}\varphi(y) \equiv Arg\min\{\varphi(y):y \in Q\}.
\end{equation}

@@

\textbf{Problem Statement A.} Find the infimum of the function $\varphi(y)$
\begin{equation}
\label{eq:1_1_5}
\varphi^*=\inf\{\varphi(y):y \in Q\}.
\end{equation}

\textbf{Problem Statement B.} Find the infimum from (\ref{eq:1_1_5}) and if the set of the global minimum points from (\ref{eq:1_1_2}) is not empty, find at least one point $y^*=Q^*$.

\textbf{Problem Statement C.} Find the infimum from (\ref{eq:1_1_5})  and all global minimum points (or make sure that the set $Q^*$ is empty).

@@

The problem statements A--D are related to the finding of the global minimum, but sometimes the following problem statement is used even for the multiextremal functions:

@@

The optimization problem statements A--E are the most commonly used (see, for example, classical monographs \cite{1_FloudasPardalos, 1_Himmelblau, 1_HorstPardalos, 1_HorstTuy, 1_McCormick, 1_Mockus, 1_Pinter,  1_SergKvasMonogr, 1_StrSergMon2000, 1_ZhigZhil}), although other variants are possible as well, for example, those given in \cite{1_Himmelblau}. The desired extremal characteristics defined by the problem statement $Z\in\{A, B, C, D, E\}$ are referred to as $Z$-solution of the optimization problem.

Let us denote the general optimization problem symbolically in the following form:
\begin{equation}
\label{eq:1_6}
\varphi(y)\rightarrow\inf, y \in Q.
\end{equation}

@@

As  the supremum of a function $\varphi(y)$ in a domain $Q$ is connected with the infimum of the function $-\varphi(y)$ according to  the expression
\begin{displaymath}
\sup\{\varphi(y):y \in Q\} = -\inf\{-\varphi(y):y \in Q\}
\end{displaymath}
the problem of finding the extremal characteristics related to the maximum of the function $\varphi(y)$ (the maximization problem), can be reduced to the problem of the minimization of the function $-\varphi(y)$. Therefore, hereinafter the optimization problem will be considered in the form (\ref{eq:1_6}). The function $\varphi(y)$ in (\ref{eq:1_6}) will be called \textit{the objective function, the minimized function, or the optimized function},  the domain $Q$ will be called \textit{the feasible (admissible) domain}, and the elements of the set $Q$ will be called \textit{the feasible points}.

Let us denote the problem (\ref{eq:1_6}), for which the set of the global minimum points $Q^*$ is known to be nonempty a priori (the sufficient conditions for the non-emptiness of $Q^*$ are given, for example, by Weierstrass theorem), as
\begin{equation}
\label{eq:1_7}
\varphi(y)\rightarrow\min, y \in Q.
\end{equation}

Note that the mathematical programming problem is often formulated just in this form, although, of course, the problem (\ref{eq:1_6}) is the more general one. Let us reformulate the problem (\ref{eq:1_7}) in the form of a non-linear programming problem
\begin{equation}
\label{eq:1_8}
\varphi(y)\rightarrow\min, y \in Q\subseteq R^N,
\end{equation}
\begin{equation}
\label{eq:1_9}
Q=\{y \in D:g_j(y)\leq 0,1\leq j\leq m \},
\end{equation}
\begin{equation}
\label{eq:1_10}
D=\{y \in R^N:y_i\in [a_i,b_i],1\leq i\leq N \},
\end{equation}
i.e., the problem of finding the extremal values (in terms of the Problem statements A-D) of the objective (minimized) function $\varphi(y)$  over the domain $Q$ is defined by the \textit{coordinate} (\ref{eq:1_10}) and \textit{functional} (\ref{eq:1_9}) \textit{constraints} on the selection of feasible points (or vectors) $y=(y_1,\ldots ,y_N)$ . In this model, the constants $a_i,b_i, 1\leq i\leq N$,  determine the limits for the variation of the problem parameters (the components of the vector $y$), and are finite or (when the corresponding upper or/and lower limits are absent for some of them) they are assumed to be equal to $a_i=-\infty $  and/or $b_i=+\infty $.

Often, the constraints in the form of equations are also included in the non-linear programming problem statement. However, any equation $h(y)=0$ , first, could be represented formally as the system of two inequalities $h(y)\leq 0$  and $h(y)\geq 0$ . Second, in numerical solving an optimization problem on a computer the exact calculation of the equality $h(y)=0$ is often impossible, therefore, this equality is assumed to be computed  with some error $\delta>0$ , i.e., the inequality $\left|h(y)\right|\leq \delta$ is considered instead of the equation $h(y)=0$. Thus, taking into account the above reasons, one can claim that (\ref{eq:1_8})--(\ref{eq:1_10}) is the statement of  \textit {the general  problem of the nonlinear programming}.

If $m=0$ , i.e., there are no functional constraints, it is considered that $Q=D$. The problem (\ref{eq:1_8})--(\ref{eq:1_10}) in this case will be called \textit {the unconstrained optimization problem}.

Let us consider an optimization problem  for the function
\begin{displaymath}
%\label{eq:1_12}
\varphi(y_1,y_2)=-1.5y_1^2\exp(1-y_1^2-20.25(y_1-y_2)^2)-
\end{displaymath}
\begin{displaymath}
%\label{eq:1_12}
-(0.5(y_1-1)(y_2-1))^4\exp(2-(0.5(y_1-1))^4-(y_2-1)^4)
\end{displaymath}
within the domain $0 \leq y_1\leq 4,\ \  -1\leq y_2\leq 3$, with the following constraints 
\begin{displaymath}
g_1(y_1,y_2)=0.01((y_1-2.2)^2+(y_2-1.2)^2-2.25)\leq 0,
\end{displaymath}
\begin{displaymath}
g_2(y_1,y_2)=100(1-(y_1-2)^2/1.44-(0.5y_2)^2)\leq 0,
\end{displaymath}
\begin{displaymath}
g_3(y_1,y_2)=10(y_2-1.5-1.5\sin(6.283(y_1-1.75)))\leq 0,
\end{displaymath}
as a particular example.

@@

The set of the points satisfying the first constraint represents the circle with the boundary defined by the equation $g_1(y_1,y_2)=0$. The points satisfying the second constraint are located outside the ellipse defined by the equation $g_2(y_1,y_2)=0$. The points satisfying the third constraint are below the sinusoid $g_3(y_1,y_2)=0$. Thus, the feasible domain is disconnect and consists of three non-convex subdomains (the feasible domain is dark in Fig.~\ref{1_fig_1}). The global minimum $\varphi(y_1^*,y_2^*)=-1.489$  is achieved at the point $(y_1^*,y_2^*)$  = (0.942, 0.944).

@@

Unfortunately, these requirements are fulfilled rarely. In the typical cases, the function $\varphi(y)$ is defined algorithmically, i.e., in the form of some calculation scheme, when one can calculate the value of $\varphi(y)$ at a given point y only (black-box function). In this case, the analytical investigation methods are unsuitable. Note that even if the function is defined analytically and if the derivative calculation is possible the initial problem (\ref{eq:1_6}) is reduced to the problem of finding all the derivative roots, the difficulty of which is comparable with the minimization problem.

@@

Let us give a formal definition of sequential optimization algorithm based on the methodology of the theory of operations research, or, more general, on the computational model for solving the problem (\ref{eq:1_6}). The generalization of this model onto the parallel computing case will be given in Subsect. \ref{subsec:1.2.2}

@@

The selection of an \textit {optimization algorithm} (method for solving  the optimization problem) is the next important step in the computational model development. In the most general form, a numerical method $s$ for solving the problem from the class $\Phi$ is a set  (tuple) (\cite{1_StrMonRus})
\begin{equation}
\label{eq:1_11}
s=\left\langle \{G_k\},\{E_k\},\{H_k\}\right\rangle
\end{equation}
where

@@

-  $\{E_k\}$	  is a set of functionals defining the rules of constructing the approximate solution (the estimates of extremum), $k=1,2,\ldots$ ;

@@

\begin{description}[\textbf{Step 1}]
\item[\textbf{Step 1}]{The point 
\begin{equation}
\label{eq:1_12}
y^1=G_1(\Phi\in Q)
\end{equation}
is selected as the first trial coordinate and the trial counter $k=1$ .}
\item[\textbf{Step 2}]{Assume that the point of the $k$-th trial $y^k\in Q$  $(k\geq 1)$ has been selected. The value of the function $z^k=\varphi(y^k)$  is computed. Afterwards there is the following \textit {search (a posteriori) information} about the function $\varphi(y)$:
\begin{equation}
\label{eq:1_13}
\omega_k=\{(y^1,z^1),(y^2,z^2),\ldots ,(y^k,z^k)\}.
\end{equation}
This information allows reducing the class, which the function  $\varphi(y)$ belongs to, up to the set
\begin{equation}
\label{eq:1_14}
\Phi(\omega_k)=\{\psi\in\Phi:\psi(y^i)=z^i,1\leq i \leq k\}.
\end{equation}
}
\item[\textbf{Step 3}]{The current estimate 
\begin{equation}
\label{eq:1_15}
e^k=E_k(\Phi,\omega_k)
\end{equation}
of the extremum sought (approximate solution) is determined.}
\item[\textbf{Step 4}]{The next trial point 
\begin{equation}
\label{eq:1_16}
y^{k+1}=G_{k+1}(\Phi,\omega_k)
\end{equation}
is computed.}
\item[\textbf{Step 5}]{The number 
\begin{equation}
\label{eq:1_17}
h^k=H_k(\Phi,\omega_k,y^{k+1})\in \{0,1\}
\end{equation}
taking one of two possible values: 0 or 1 is determined. If $h^k=1$ the trial counter $k$ is increased by 1, and the return to the Step 2 of the scheme is carried out. If $h^k=0$ the computations are terminated, and the estimate $e^k$  is taken as the problem solution.}
\end{description}

@@

\example {Let us consider the simplest method of solving a one-dimensional problem (\ref{eq:1_6}) within an interval $[a,b]$, namely, the method of scanning  the function values in the nodes of a regular grid. The method consists in the partitioning of the search interval $[a,b]$ into $q$ equal parts. The function values are calculated at the points (nodes) of the partitioning, including the ends of the search interval. The lowest calculated value (and its coordinate, if it is required by the problem statement) is considered to be the solution of the problem. }

The computability of the function at any point of the search domain is sufficient for the applicability of the method. So, the class of the functions defined over the interval $[a,b]$ and being computable at every point of the interval can be considered as the a priori known class $\Phi$.

For this method
\begin{displaymath}
G_1(\Phi)=a, G_{k+1}(\Phi,\omega_k)=a+k\frac{b-a}{q}, k\geq 1,
\end{displaymath}
\begin{displaymath}
  H_k(\Phi,\omega_k) =
  \begin{cases}
    0, & k > q, \\
    1, & k \leq q,
  \end{cases}
\end{displaymath}
\begin{equation}
\label{eq:1_18}
e^k=\varphi_k^*,
\end{equation}
where
\begin{equation}
\label{eq:1_19}
\varphi_k^*=\min_{1\leq i\leq k} \varphi(y^i),
\end{equation}
or
\begin{displaymath}
e^k=(\varphi_k^*,y_k^*),
\end{displaymath}
where 
\begin{displaymath}
%\label{eq:1_19}
y_k^*=\arg \min_{1\leq i\leq k} \varphi(y^i).
\end{displaymath}

\subsection {Parallel optimization algorithms}
\label {subsec:1.2.2}
The computational scheme (\ref{eq:1_12})--(\ref{eq:1_17}) implies  the sequential execution of all the algorithm steps, that can be provided by the single processor computers. But if a multiprocessor computer is available, a parallelization of the computational scheme and, therefore, an acceleration of the optimization process can be realized. The parallelization of optimization methods could be done in several ways. It is possible to parallelize, in the first place, the calculation of the objective function, secondly, the implementation of the rules $G_k,E_k,H_k$ , and, finally, the scheme (\ref{eq:1_12})--(\ref{eq:1_17}) providing simultaneous execution of several trials. However, the parallelization of the objective function describing the optimized object is specific for different objective function models, and as a matter of fact, cannot be related to the general optimization scheme. The parallelization of the functionals  $G_k,E_k,H_k$ depends essentially on the class of algorithms. Besides, often these functionals are rather simple, and their parallelization is just of no use. Therefore, the subject of the consideration in the present subsection  is the consideration of some \textit {fundamental principles} for a parallel execution of trials.  Let us use the term \textit {iteration} for the simultaneous (parallel) execution of several trials (each trial is executed by a separate processor). Let us denote the number of trials during the $n$-th iteration as $p(n)$ , and the total number of trials executed during all $n$ iterations as $k(n)$ . In other words, we assume that we have  $p$  processors at our disposal while executing the $n$-th iteration. Actually, the number of processors employed can be different for different iterations.

@@

Thereby, let us ask the question: what requirements should be imposed to trial sequences of a numerical optimization method? Of course, the main requirement is that execution of the trials at the points $\{y^k\}$  with the trial results $\{z^k\}$  could provide a solution of the problem, i.e., finding the solution corresponding to the selected problem statement. At the same time, since a computer can execute a finite number of trials only, it is desirable to obtain the exact solution after constructing a finite sequence $\{y^k\}$. Unfortunately, such a favorable situation takes place in rare and simple cases only, for example, in the linear programming problems. Therefore,  an asymptotically exact assessment is interesting often, and an infinite trial sequence is considered (in this case for the introduced  optimization method models in (\ref{eq:1_17}) , (\ref{eq:1_25}), (\ref{eq:1_30}) $H_k=1$  for any $k\geq1$, i.e., in fact, the termination criterion is absent). In other words, this sequence is required to converge to the exact problem solution. Since in the problem statements $B-D$ the desired solution may include several minimum points, we will understand the convergence of a method in the sense of the following definition.
\begin{definition} 
\label{def:1_4}
The trial sequence $\{y^k\}$  converges to the solution of the problem (\ref{eq:1_6}) defined by the corresponding problem statement, if:
\begin{enumerate}
\item{it includes a subsequence  $\{\bar{y}^k\}$, for which $\lim_{k \to \infty}\bar{y}^k=\varphi^*$;}
\item{in the case when the solution includes one or several minimum points, for each such point there exists a subsequence of the sequence $\{y^k\}$  converging to it.}
\end{enumerate}
\end{definition}
Trial sequence converging to the exact solution of the problem statement $Z \in \{A,B,C,D,E\}$ is called the minimizing sequence for the problem statement $Z$ or $Z$-minimizing sequence. The term «minimizing sequence» was introduced in \cite{1_Vasiliev} and corresponds to the concept of $A$-minimizing sequence.

@@

Another approach to the design of the numerical methods for the analysis of the multidimensional multiextremal problems exploits the idea of \textit{the complexity reduction} where the solution of the initial problem is replaced with solving one or several simpler problems.
